---
title: "AMP_ves_summary_plots"
author: "Jessica McCordic"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

```{r load libraries and helper script, include=FALSE, message=FALSE, warning=FALSE, echo=FALSE}

#### Load libraries ####

# NOAA-approved tidyverse
tidyverse_short<-c("broom","cli","crayon","dbplyr","dplyr","dtplyr","forcats","ggplot2","googledrive","googlesheets4","hms","httr","jsonlite","lubridate","magrittr","modelr","pillar","purrr","readr","readxl","reprex","rlang","rstudioapi","rvest","stringr","tibble","tidyr","xml2") 
lapply(tidyverse_short, require, character.only = TRUE) 

library(scales)
library(ggalt)
library(ggsci)
library(ggdist)
library(ggh4x) # nested faceting

theme_set(theme_bw(
  base_size = 14
))

# source helper file
source("code/AMP_summary_ves_funs.R")
```

## Load hourly presence datasets

```{r Load data, include=FALSE}

#### Load compiled hourly presence table ####
# This is generated using Reformat_data_for_HP script

all_sites_hp <- read_csv("data/All_sites_tables/hourly_pres_allsites_local.csv")
all_sites_inout <- read_csv("data/All_sites_tables/total_ves_ins_out_by_site.csv")

```

## Reshape data for plotting

```{r reshape tables, include = FALSE}

#### Add columns to original dataset to help group for plotting ####

all_sites_hp <- all_sites_hp |>
  mutate(
    # replace NA's with 0's to accurately incorporate hours with presence = 0
    Total_Vessels = replace_na(Total_Vessels, 0),
    total_inside = replace_na(total_inside, 0),
    total_inside_small = replace_na(total_inside_small, 0),
    
    # add categorical Y/N column for vessels present to calculate proportions
    ves_yn = ifelse(Total_Vessels == 0, "N","Y"),
    ins_ves_yn = ifelse(total_inside == 0, "N","Y"),
    ins_ves_small_yn = ifelse(total_inside_small == 0, "N","Y"),
    
    # add date without year to collapse across years
    Month = month(Begin_Date_loc),
    Week = isoweek(Begin_Date_loc),
    Day_month = mday(Begin_Date_loc),
    # re-create dates as all in the same year for plotting on single graph
    Date_noyr = as_date(paste0("2000", "-", Month, "-", Day_month), format = "%Y-%m-%d"),
    # get start date for each week
    Date_week = as_date(paste(as.character(year(Begin_Date_loc)),Week,"1",sep = "-"), format = "%Y-%W-%w"),
    
    # add season based on month
    season = case_when(month(Date_noyr) == 12 | month(Date_noyr) == 1 | month(Date_noyr) == 2 ~ "Summer",
                       month(Date_noyr) == 3 | month(Date_noyr) == 4 | month(Date_noyr) == 5 ~ "Autumn",
                       month(Date_noyr) == 6 | month(Date_noyr) == 7 | month(Date_noyr) == 8 ~ "Winter",
                        TRUE ~ "Spring"
    ),
    
    # add AMP Regional Networks and NPZ IDs based on Site_ID
    network = case_when(Site_ID == "CG" | Site_ID == "EP" | Site_ID == "WP" ~ "Temperate East",
                        Site_ID == "DNE" | Site_ID == "DSW" | Site_ID == "NGN"| Site_ID == "NGS" ~ "Northwest",
                        TRUE ~ "South-west"
    ),
    npz_id = case_when(
      # Cod Grounds
      Site_ID == "CG" ~ "CGMP", 
      # Solitary Islands
      Site_ID == "EP" | Site_ID == "WP" ~ "SIMP",
      # Dampier
      Site_ID == "DNE" | Site_ID == "DSW" ~ "DMP",
      # Ningaloo
      Site_ID == "NGN"| Site_ID == "NGS" ~ "NMP", 
      # Two Rocks
      Site_ID == "TRE" | Site_ID == "TRW" ~ "TRMP", 
      # Jurien
      Site_ID == "JSE" | Site_ID == "JNE" ~ "JMP", 
      # Geographe
      Site_ID == "GEO" ~ "GMP",
      # Murat
      Site_ID == "MRE" | Site_ID == "MRW" ~ "MMP",
      # South-west Corner
      Site_ID == "SWS" ~ "SWCMP" 
    )
  )

```

## Data gaps

Add in dates without coverage 
**(this is still in progress)**

```{r data gaps, include=FALSE}

##### Deployment info ###

# create table of data gaps within nominal year (2000)

gaps <- data.frame("Gap_start" = c(as.POSIXct("2000-02-21",format="%Y-%m-%d"), #CGMP gap 1 start
                                   as.POSIXct("2000-08-05",format="%Y-%m-%d"), #CGMP gap 2 start
                                   as.POSIXct("2000-01-16",format="%Y-%m-%d"), #SIMP gap 1 start
                                   as.POSIXct("2000-09-12",format="%Y-%m-%d"), #SIMP gap 2 start
                                   as.POSIXct("2000-01-01",format="%Y-%m-%d"), #DMP gap 1 start
                                   as.POSIXct("2000-11-05",format="%Y-%m-%d"), #DMP gap 2 start
                                   as.POSIXct("2000-01-01",format="%Y-%m-%d"), #NMP gap 1 start
                                   as.POSIXct("2000-11-19",format="%Y-%m-%d"), #NMP gap 2 start
                                   as.POSIXct("2000-01-01",format="%Y-%m-%d"), #GMP gap 1 start
                                   as.POSIXct("2000-04-04",format="%Y-%m-%d"), #GMP gap 2 start
                                   as.POSIXct("2000-01-01",format="%Y-%m-%d"), #JMP gap 1 start
                                   as.POSIXct("2000-03-11",format="%Y-%m-%d"), #JMP gap 2 start
                                   as.POSIXct("2000-01-01",format="%Y-%m-%d"), #MMP gap 1 start
                                   as.POSIXct("2000-05-04",format="%Y-%m-%d"), #MMP gap 2 start
                                   as.POSIXct("2000-01-01",format="%Y-%m-%d"), #SWCMP gap 1 start
                                   as.POSIXct("2000-04-12",format="%Y-%m-%d"), #SWCMP gap 2 start
                                   as.POSIXct("2000-01-01",format="%Y-%m-%d"), #TRMP gap 1 start
                                   as.POSIXct("2000-08-03",format="%Y-%m-%d") #TRMP gap 2 start
                                   ),
                                    
                   
                   "Gap_end" = c(as.POSIXct("2000-04-10",format="%Y-%m-%d"), #CGMP gap 1 end
                                   as.POSIXct("2000-12-11",format="%Y-%m-%d"), #CGMP gap 2 end
                                   as.POSIXct("2000-06-09",format="%Y-%m-%d"), #SIMP gap 1 end
                                   as.POSIXct("2000-12-11",format="%Y-%m-%d"), #SIMP gap 2 end
                                   as.POSIXct("2000-09-08",format="%Y-%m-%d"), #DMP gap 1 end
                                   as.POSIXct("2000-12-31",format="%Y-%m-%d"), #DMP gap 2 end
                                   as.POSIXct("2000-09-24",format="%Y-%m-%d"), #NMP gap 1 end
                                   as.POSIXct("2000-12-31",format="%Y-%m-%d"), #NMP gap 2 end
                                   as.POSIXct("2000-01-12",format="%Y-%m-%d"), #GMP gap 1 end
                                   as.POSIXct("2000-12-31",format="%Y-%m-%d"), #GMP gap 2 end
                                   as.POSIXct("2000-01-04",format="%Y-%m-%d"), #JMP gap 1 end
                                   as.POSIXct("2000-12-31",format="%Y-%m-%d"), #JMP gap 2 end
                                   as.POSIXct("2000-02-07",format="%Y-%m-%d"), #MMP gap 1 end
                                   as.POSIXct("2000-12-31",format="%Y-%m-%d"), #MMP gap 2 end
                                   as.POSIXct("2000-01-07",format="%Y-%m-%d"), #SWCMP gap 1 end
                                   as.POSIXct("2000-12-31",format="%Y-%m-%d"), #SWCMP gap 2 end
                                   as.POSIXct("2000-05-09",format="%Y-%m-%d"), #TRMP gap 1 end
                                   as.POSIXct("2000-12-31",format="%Y-%m-%d") #TRMP gap 2 end
                                 ),
                   
                   "npz_id" = c("CGMP",
                              "CGMP",
                              "SIMP",
                              "SIMP",
                              "DMP",
                              "DMP",
                              "NMP",
                              "NMP",
                              "GMP",
                              "GMP",
                              "JMP",
                              "JMP",
                              "MMP",
                              "MMP",
                              "SWCMP",
                              "SWCMP",
                              "TRMP",
                              "TRMP"))


gaps <- gaps |>
  mutate(network = case_when(npz_id == "CGMP" | npz_id == "SIMP" ~ "Temperate East",
                             npz_id == "DMP" | npz_id == "NMP" ~ "Northwest", TRUE ~ "South-west"
         ),
         network = factor(network, levels = c("Temperate East","Northwest","South-west")))



gaps_wk <- gaps |>
  mutate(Gap_start = floor_date(Gap_start, "1 week"),
         Gap_end = floor_date(Gap_end, "1 week"))

```

## Data summary and visualization

### Total vessel presence per deployment

Raw total counts per deployments and vessels inside the NPZs
Deployment-level and NPZ-level info

```{r total presence, message=FALSE, warning=FALSE}

# summarize in-out data to get n vessels used for I-O analysis per deployment
ins_out_counts <- all_sites_inout |>
  group_by(Dep_ID) |>
  summarize(n_ves_used_inout = n()) |>
  rename("Dep" = "Dep_ID")

# summarize full df to get counts per deployment
all_ves_by_site <- all_sites_hp |>
  group_by(network, npz_id, Site_ID, Dep_ID, Dep) |>
  summarise(n_days = n_distinct(Begin_Date_loc),
            Total_ves_dep = sum(Total_Vessels, na.rm = TRUE),
            Total_inside = sum(total_inside, na.rm = TRUE),
            Total_inside_small = sum(total_inside_small, na.rm = TRUE),
            Inside_maneuver = sum(man_inside, na.rm = TRUE),
            Inside_maneuver_small = sum(man_inside_small, na.rm = TRUE),
            mean_per_day = Total_ves_dep/n_days,
            mean_overall = mean(Total_Vessels)) |>
  mutate(Inside_transit = Total_inside - Inside_maneuver,
         Inside_transit_small = Total_inside_small - Inside_maneuver_small) |>
  left_join(ins_out_counts, by = "Dep") |>
  mutate(
    prop_inside = Total_inside/n_ves_used_inout,
    total_outside = n_ves_used_inout - Total_inside,
    prop_inside_small = Total_inside_small/n_ves_used_inout,
    total_out_small = n_ves_used_inout - Total_inside_small)


all_ves_by_npz <- all_sites_hp |>
  group_by(network, npz_id, Site_ID, Dep_ID, Dep) |>
  left_join(ins_out_counts, by = "Dep") |>
  ungroup() |>
  group_by(network, npz_id) |>
  summarise(n_days = n_distinct(Site_ID, Begin_Date_loc),
            Total_ves_npz = sum(Total_Vessels, na.rm = TRUE),
            Total_inside = sum(total_inside, na.rm = TRUE),
            Total_inside_small = sum(total_inside_small, na.rm = TRUE),
            
            Inside_maneuver = sum(man_inside, na.rm = TRUE),
            
            Inside_maneuver_small = sum(man_inside_small, na.rm = TRUE),
            mean_per_day = Total_ves_npz/n_days,
            
            n_used_inout = sum(unique(n_ves_used_inout)),
            
            mean_overall = mean(Total_Vessels),
           ) |>
  mutate(Inside_transit = Total_inside - Inside_maneuver,
         Inside_transit_small = Total_inside_small - Inside_maneuver_small,
         total_outside = n_used_inout - Total_inside,
         total_out_small = n_used_inout - Total_inside_small,
         prop_inside = Total_inside/n_used_inout,
         prop_inside_small = Total_inside_small/sum(n_used_inout))



all_ves_long <- all_ves_by_site |>
  pivot_longer(cols = -c(network, npz_id, Site_ID, Dep_ID, Dep, n_days), values_to = "Ves_counts", names_to = "Ves_behav") |>
  # relevel categorical values for plotting
  mutate(Ves_behav = fct(Ves_behav),
         Ves_behav = fct_relevel(Ves_behav, c("Total_ves_dep","Total_inside_small", "Total_inside", "Inside_transit_small","Inside_maneuver_small", "Inside_transit","Inside_maneuver")),
         network = fct(network),
         network = fct_relevel(network, c("Temperate East","Northwest"))) |>
  # add diverging counts where inside values are negative for plotting
  mutate(ves_counts_diverge = ifelse(Ves_behav == "Total_inside" | Ves_behav == "Total_inside_small",Ves_counts*(-1),Ves_counts),
         npz_id = fct(npz_id),
         npz_id = fct_relevel(npz_id, c("CGMP","SIMP","DMP","NMP")))


all_ves_by_day <- all_sites_hp |>
  group_by(network, npz_id, Site_ID, Dep_ID, Dep, Date_week, Begin_Date_loc) |>
  summarise(n_hrs = n_distinct(Begin_Hour_loc),
            n_ves_hrs = sum(ifelse(ves_yn == "Y", 1, 0), na.rm = TRUE),
            Total_ves_dep = sum(Total_Vessels, na.rm = TRUE),
            Total_inside = sum(total_inside, na.rm = TRUE),
            Total_inside_small = sum(total_inside_small, na.rm = TRUE),
            Inside_maneuver = sum(man_inside, na.rm = TRUE),
            Inside_maneuver_small = sum(man_inside_small, na.rm = TRUE),
            # daily means
            mean_total_ves = mean(Total_Vessels),
            mean_ins_ves = mean(total_inside),
            mean_ins_sm_ves = mean(total_inside_small)) |>
  mutate(daily_ves_perhr = Total_ves_dep/n_hrs,
         total_prophr = n_ves_hrs/n_hrs,
        Inside_transit = Total_inside - Inside_maneuver,
         Inside_transit_small = Total_inside_small - Inside_maneuver_small)

all_ves_by_day_noyr <- all_sites_hp |>
  group_by(network, npz_id, Site_ID, Dep_ID, Dep, Date_week, Date_noyr) |>
  summarise(n_hrs = n_distinct(Begin_Hour_loc),
            n_ves_hrs = sum(ifelse(ves_yn == "Y", 1, 0), na.rm = TRUE),
            Total_ves_dep = sum(Total_Vessels, na.rm = TRUE),
            Total_inside = sum(total_inside, na.rm = TRUE),
            Total_inside_small = sum(total_inside_small, na.rm = TRUE),
            Inside_maneuver = sum(man_inside, na.rm = TRUE),
            Inside_maneuver_small = sum(man_inside_small, na.rm = TRUE),
            # daily means
            mean_total_ves = mean(Total_Vessels),
            mean_ins_ves = mean(total_inside),
            mean_ins_sm_ves = mean(total_inside_small)) |>
  mutate(daily_ves_perhr = Total_ves_dep/n_hrs,
         total_prophr = n_ves_hrs/n_hrs,
         Inside_transit = Total_inside - Inside_maneuver,
         Inside_transit_small = Total_inside_small - Inside_maneuver_small,
         network = case_when(npz_id == "CGMP" | npz_id == "SIMP" ~ "Temperate East",
                             npz_id == "DMP" | npz_id == "NMP" ~ "Northwest", TRUE ~ "South-west"
         ),
         network = factor(network, levels = c("Temperate East","Northwest","South-west"))) 



```

### Plots by NPZ

Visualizing overall vessel presence

```{r plots by npz}
ggplot(data = all_ves_by_site,
       aes(x = npz_id,
           y = Total_ves_dep,
           fill = npz_id)) +
  geom_col(show.legend = FALSE,
           position = position_identity()) +
  scale_fill_viridis_d()+
  facet_grid(~network, 
             scales = "free_x", 
             space = "free", 
             drop = TRUE) +
  labs(x = "NPZ",
       y = "N vessels",
       title = "Total vessel counts") +
  facet_nested(cols = vars(network, npz_id, Dep_ID), 
               nest_line = element_line(),
               scales = "free_x") +
  theme(
    axis.text.x = element_text(face = "bold"),
    strip.background = element_blank(),
    strip.text.y = element_text(face = "bold")
    )

# weekly plots for nominal year
ggplot() +
  geom_boxplot(data = all_ves_by_day_noyr |> filter(total_prophr > 0),
       aes(x = as_date(Date_noyr),
           y = total_prophr, 
           group = as_date(Date_week),
           fill = network),
               show.legend = FALSE) + 
    geom_rect(data = gaps,
            aes(xmin = as_date(Gap_start),
                xmax = as_date(Gap_end),
                ymin = 0,
                ymax = 0.8),
            fill = "grey80",
            alpha = 0.7)+
  scale_x_date(date_breaks = "months", 
               date_labels = "%b")+
scale_fill_manual(values = c("Temperate East" = "#00AB8E",
                             "Northwest" = "#A05EB5",
                             "South-west" = "#485CC7"))+
  labs(y = "N hours present /\nN total hours",
       x = "Date")+
  facet_nested(rows = vars(network, npz_id), 
               nest_line = element_line(),
               scales = "free_x") +
    theme(
    axis.text.x = element_text(face = "bold"),
    strip.background = element_blank(),
    strip.text.y = element_text(face = "bold")
    )

ggsave(paste0("figs/", "Total_boxplot_daily_prophr_1yr.png"), width=10, height=8,
       units="in", dpi=300)

```


### Diverging bar plot by NPZ

Of vessels suitable for propagation analysis, what proportion are estimated to occur within the NPZ boundaries?

```{r div bar by npz}

####
# Diverging bar plot - NPZ
#### 


breakvals <- seq(-200, 800, 200)

ggplot(data = all_ves_long |> 
         filter(Ves_behav == "total_outside" |
               Ves_behav == "Total_inside",
               npz_id != "SWCMP"),
       aes(x = npz_id,
           y = ves_counts_diverge,
           fill = Ves_behav)) +
  geom_col(position = position_stack(reverse = TRUE)) +
  coord_flip()+
  scale_y_continuous(limits = c(-200, 800),
                     breaks = breakvals, 
                     labels = abs(breakvals))+
  scale_fill_viridis_d(direction = -1)+
  labs(y = "N vessels",
       fill = "Inside vs. Outside",
       title = "Inside-Outside Totals") +
  facet_nested(rows = vars(network, npz_id, Dep_ID), 
               nest_line = element_line(),
               scales = "free_y",
               switch = "y") +
  theme(
    axis.text.x = element_text(face = "bold"),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    strip.background = element_blank(),
    strip.text.y = element_text(face = "bold")
    )
  
ggsave(paste0("figs/", "Ins_out_counts_npz.png"), width=10, height=8,
       units="in", dpi=300)

# Small vessels inside-outside

breakvals_small <- seq(-600, 600, 200)

ggplot(data = all_ves_long |> 
         filter(Ves_behav == "total_out_small" |
               Ves_behav == "Total_inside_small",
               npz_id != "SWCMP"),
       aes(x = npz_id,
           y = ves_counts_diverge,
           fill = Ves_behav)) +
  geom_col(position = position_stack(reverse = TRUE)) +
  coord_flip()+
  scale_y_continuous(limits = c(-600, 600),
                     breaks = breakvals_small, 
                     labels = abs(breakvals_small))+
  scale_fill_viridis_d(direction = -1)+
  labs(y = "N vessels",
       fill = "Inside vs. Outside",
       title = "Inside-Outside Counts (Small Vessels)") +
   facet_nested(rows = vars(network, npz_id, Dep_ID), 
               nest_line = element_line(),
               scales = "free_y",
               switch = "y") +
  theme(
    axis.text.x = element_text(face = "bold"),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    strip.background = element_blank(),
    strip.text.y = element_text(face = "bold")
    )

ggsave(paste0("figs/", "Ins_out_counts_small_npz.png"), width=10, height=8,
       units="in", dpi=300)


# make table of inside vs. outside
ins_out_table <- all_ves_long |> 
  filter(Ves_behav == "total_outside" |
           Ves_behav == "Total_inside"| 
           Ves_behav == "total_out_small" |
           Ves_behav == "Total_inside_small",
         npz_id != "SWCMP") |>
  pivot_wider(names_from = Ves_behav, 
              id_cols = Dep, 
              values_from = Ves_counts) |>
  relocate(total_outside, .after = Total_inside) |>
  rename("Total inside" = "Total_inside",
         "Total inside (small)" = "Total_inside_small",
         "Total outside" = "total_outside",
         "Total outside (small)" = "total_out_small",
         "Deployment" = "Dep")

knitr::kable(ins_out_table, caption = "Vessel counts inside and outside NPZ boundaries")



```



## Temporal patterns

### Weekly mean-adjusted weekday presence

Calculate mean-adjusted totals per weekday, similar to Moussoline et al. 2012 diel presence

- For all full weeks (7d), Calculate weekly mean of daily totals per deployment
- For each weekday, calculate [daily total - weekly mean]
- values will be +/- relative to weekly mean

For weeks with vessel presence, the mean-adjusted weekday counts were highest in ___ NPZ on __day. Multiple NPZs showed higher presence on __days overall. 


```{r weekday mean adjusted}

# start by getting daily counts
all_ves_wkdy <- all_sites_hp |>
  group_by(network, npz_id, Dep_ID, Dep, Date_week, Weekday, Begin_Date_loc)|>
  # get daily vessel counts per weekday: total, inside, inside_small, maneuver, maneuver_small
  summarise(Total_ves_dep = sum(Total_Vessels, na.rm = TRUE),
            Total_inside = sum(total_inside, na.rm = TRUE),
            Total_inside_small = sum(total_inside_small, na.rm = TRUE),
            Inside_maneuver = sum(man_inside, na.rm = TRUE),
            Inside_maneuver_small = sum(man_inside_small, na.rm = TRUE)) |>
  mutate(Inside_transit = Total_inside - Inside_maneuver,
         Inside_transit_small = Total_inside_small - Inside_maneuver_small, 
         Weekday = factor(Weekday, levels=c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday")),
         network = case_when(npz_id == "CGMP" | npz_id == "SIMP" ~ "Temperate East",
                             npz_id == "DMP" | npz_id == "NMP" ~ "Northwest", TRUE ~ "South-west"
         ),
         network = factor(network, levels = c("Temperate East","Northwest","South-west"))) 
levels(all_ves_wkdy$Weekday) <- c("M","T","W","R","F","Sa","Su")


# Weekly mean-adjusted counts for total vessels
wkly_means_tot <- all_ves_wkdy |>
  # group to week level
  group_by(network, npz_id, Dep_ID, Dep, Date_week) |>
  # get n_days per week in dataset
  summarize(n_days = n(),
            week_sum = sum(Total_ves_dep, na.rm = TRUE)) |>
  # only keep full weeks n_days == 7
  # only keep weeks with presence
  filter(n_days == 7,
         week_sum > 0) |>
  # get original weekday counts
  left_join(all_ves_wkdy) |>
  # re-group to week level
  group_by(network, npz_id, Dep_ID, Dep, Date_week) |>
  # calculate weekly mean of total presence for days with presence
  mutate(mean_tot = mean(Total_ves_dep, na.rm = TRUE),
         # calculate mean-adjusted counts per weekday
         tot_wkdy_adj = Total_ves_dep - mean_tot) |>
  group_by(network, npz_id, Dep_ID, Weekday) |>
  # get mean of mean-adjusted counts for each weekday
  summarise(mean_wkdy_adj = mean(tot_wkdy_adj, na.rm = TRUE)) |>
  # get lower & upper bounds of SD of the mean of the adjusted counts
  mutate(sd_lower = mean_wkdy_adj - 0.5*(sd(mean_wkdy_adj)),
         sd_upper = mean_wkdy_adj + 0.5*(sd(mean_wkdy_adj)))




# Weekly mean-adjusted counts, inside and inside maneuvers
wkly_means_ins <- all_ves_wkdy |>
  # group to week level
  group_by(network, npz_id, Dep_ID, Dep, Date_week) |>
  # get n_days per week in dataset
  summarize(n_days = n(),
            week_sum = sum(Total_inside, na.rm = TRUE)) |>
  # only keep full weeks n_days == 7
  # only keep weeks with presence
  filter(n_days == 7,
         week_sum > 0) |>
  # get original weekday counts
  left_join(all_ves_wkdy) |>
  # re-group to week level
  group_by(network, npz_id, Dep_ID, Dep, Date_week) |>
  # calculate weekly mean of total presence for days with presence
  mutate(mean_tot = mean(Total_inside, na.rm = TRUE),
         mean_man = mean(Inside_maneuver, na.rm = TRUE),
         # calculate mean-adjusted counts per weekday
         tot_wkdy_adj = Total_inside - mean_tot,
         man_wkdy_adj = Inside_maneuver - mean_man) |>
  group_by(network, npz_id, Dep_ID, Weekday) |>
  # get mean of mean-adjusted counts for each weekday
  summarise(mean_wkdy_adj = mean(tot_wkdy_adj, na.rm = TRUE),
            mean_man_wkdy_adj = mean(man_wkdy_adj, na.rm = TRUE)) |>
  # get lower & upper bounds of SD of the mean of the adjusted counts
  mutate(sd_lower = mean_wkdy_adj - 0.5*(sd(mean_wkdy_adj)),
         sd_upper = mean_wkdy_adj + 0.5*(sd(mean_wkdy_adj)),
         man_sd_lower = mean_man_wkdy_adj - 0.5*(sd(mean_man_wkdy_adj)),
         man_sd_upper = mean_man_wkdy_adj + 0.5*(sd(mean_man_wkdy_adj)))


# Weekly mean-adjusted counts, inside and inside maneuvers
wkly_means_ins_man <- all_ves_wkdy |>
  # group to week level
  group_by(network, npz_id, Dep_ID, Dep, Date_week) |>
  # get n_days per week in dataset
  summarize(n_days = n(),
            week_sum = sum(Inside_maneuver, na.rm = TRUE)) |>
  # only keep full weeks n_days == 7
  # only keep weeks with presence
  filter(n_days == 7,
         week_sum > 0) |>
  # get original weekday counts
  left_join(all_ves_wkdy) |>
  # re-group to week level
  group_by(network, npz_id, Dep_ID, Dep, Date_week) |>
  # calculate weekly mean of total presence for days with presence
  mutate(mean_man = mean(Inside_maneuver, na.rm = TRUE),
         # calculate mean-adjusted counts per weekday
         man_wkdy_adj = Inside_maneuver - mean_man) |>
  group_by(network, npz_id, Dep_ID, Weekday) |>
  # get mean of mean-adjusted counts for each weekday
  summarise(mean_man_wkdy_adj = mean(man_wkdy_adj, na.rm = TRUE)) |>
  # get lower & upper bounds of SD of the mean of the adjusted counts
  mutate(man_sd_lower = mean_man_wkdy_adj - 0.5*(sd(mean_man_wkdy_adj)),
         man_sd_upper = mean_man_wkdy_adj + 0.5*(sd(mean_man_wkdy_adj)))


# Weekly mean-adjusted counts, inside small only
wkly_means_ins_small <- all_ves_wkdy |>
  # group to week level
  group_by(network, npz_id, Dep_ID, Dep, Date_week) |>
  # get n_days per week in dataset
  summarize(n_days = n(),
            week_sum = sum(Total_inside_small, na.rm = TRUE)) |>
  # only keep full weeks n_days == 7
  # only keep weeks with presence
  filter(n_days == 7,
         week_sum > 0) |>
  # get original weekday counts
  left_join(all_ves_wkdy) |>
  # re-group to week level
  group_by(network, npz_id, Dep_ID, Dep, Date_week) |>
  # calculate weekly mean of total presence for days with presence
  mutate(mean_tot = mean(Total_inside_small, na.rm = TRUE),
         # calculate mean-adjusted counts per weekday
         tot_wkdy_adj = Total_inside_small - mean_tot) |>
  group_by(network, npz_id, Dep_ID, Weekday) |>
  # get mean of mean-adjusted counts for each weekday
  summarise(mean_wkdy_adj = mean(tot_wkdy_adj, na.rm = TRUE)) |>
  # get lower & upper bounds of SD of the mean of the adjusted counts
  mutate(sd_lower = mean_wkdy_adj - 0.5*(sd(mean_wkdy_adj)),
         sd_upper = mean_wkdy_adj + 0.5*(sd(mean_wkdy_adj)))

```



### Plot mean-adjusted weekday counts

```{r plot weekly mean adj, echo=FALSE}
ggplot(data = wkly_means_tot,
       aes(x = Weekday, y = mean_wkdy_adj)) +
  geom_line(aes(group = Dep_ID),
            color = "gray70")+
  geom_pointrange(aes(ymin = sd_lower,
                      ymax = sd_upper,
                      group = npz_id,
                      color = ifelse(sd_lower <= 0, 'firebrick','blue')),
                  size = 0.25,
                  show.legend = FALSE) +
  geom_hline(yintercept = 0, 
             linetype = 3, # dotted line
             colour = "black") +
  scale_color_manual(values = c("red","black"))+
  labs(title = "Total vessels",
       y = "Mean-adjusted average N vessels / weekday",
       x = "Weekday")+
  facet_nested(rows = vars(network, npz_id, Dep_ID), 
               nest_line = element_line(),
               scales = "free_y") +
  theme(
    axis.text.x = element_text(face = "bold"),
    strip.background = element_blank(),
    strip.text.y = element_text(face = "bold")
    )

ggsave(paste0("figs/", "Weekday_mean_adj_dep_presentonly.png"), width=5, height=12, units="in", dpi=300)


#### Try plotting total & inside on same plots...? ####
# not necessarily for publication, just helps me wrap my head around patterns

ggplot(data = wkly_means_tot,
       aes(x = Weekday, y = mean_wkdy_adj)) +
  geom_hline(yintercept = 0, 
             linetype = 3, # dotted line
             colour = "black") +
  geom_line(aes(group = Dep_ID),
            color = "black")+
  geom_pointrange(aes(ymin = sd_lower,
                      ymax = sd_upper,
                      group = npz_id,
                      color = ifelse(sd_lower <= 0, 'gray70','black')),
                  size = 0.25,
                  show.legend = FALSE) +
  geom_line(data = wkly_means_ins,
       aes(x = Weekday, y = mean_wkdy_adj, group = Dep_ID),
            color = "blue")+
  geom_pointrange(data = wkly_means_ins,
                  aes(ymin = sd_lower,
                      ymax = sd_upper,
                      group = npz_id,
                      color = ifelse(sd_lower <= 0, 'lightblue','darkblue')),
                  size = 0.25,
                  show.legend = FALSE) +
  
  scale_color_manual(values = c('red', # above 0, total
                                'darkblue', # above 0, inside
                                'black', # below 0, total
                                'gray70' # below 0, inside
                                ))+
  labs(title = "Total vessels",
       y = "Mean-adjusted average N vessels / weekday",
       x = "Weekday")+
  facet_nested(rows = vars(network, npz_id, Dep_ID), 
               nest_line = element_line(),
               scales = "free_y") +
  theme(
    axis.text.x = element_text(face = "bold"),
    strip.background = element_blank(),
    strip.text.y = element_text(face = "bold")
    )


ggsave(paste0("figs/", "Weekday_mean_adj_combined_total_ins.png"), width=5, height=12, units="in", dpi=300)







# Inside vessels
ggplot(data = wkly_means_ins,
       aes(x = Weekday, y = mean_wkdy_adj)) +
  geom_line(aes(group = Dep_ID),
            color = "gray70")+
  geom_pointrange(aes(ymin = sd_lower,
                      ymax = sd_upper,
                      group = npz_id,
                      color = ifelse(sd_lower <= 0, 'firebrick','blue')),
                  size = 0.25,
                  show.legend = FALSE) +
  geom_hline(yintercept = 0, 
             linetype = 3, # dotted line
             colour = "black") +
  scale_color_manual(values = c("red","black"))+
  labs(title = "Inside vessels",
    y = "Mean-adjusted average N vessels / weekday",
       x = "Weekday")+
  facet_nested(rows = vars(network, npz_id, Dep_ID), 
               nest_line = element_line(),
               scales = "free_y") +
  theme(
    axis.text.x = element_text(face = "bold"),
    strip.background = element_blank(),
    strip.text.y = element_text(face = "bold")
    )
ggsave(paste0("figs/", "Weekday_mean_adj_dep_presentonly_inside.png"), width=5, height=12, units="in", dpi=300)


# Inside maneuvers only
# Inside vessels
ggplot(data = wkly_means_ins_man,
       aes(x = Weekday, y = mean_man_wkdy_adj)) +
  geom_line(aes(group = Dep_ID),
            color = "gray70")+
  geom_pointrange(aes(ymin = man_sd_lower,
                      ymax = man_sd_upper,
                      group = npz_id,
                      color = ifelse(man_sd_lower <= 0, 'firebrick','blue')),
                  size = 0.25,
                  show.legend = FALSE) +
  geom_hline(yintercept = 0, 
             linetype = 3, # dotted line
             colour = "black") +
  scale_color_manual(values = c("red","black"))+
  labs(title = "Inside maneuvers",
    y = "Mean-adjusted average N vessels / weekday",
       x = "Weekday")+
  facet_nested(rows = vars(network, npz_id, Dep_ID), 
               nest_line = element_line(),
               scales = "free_y") +
  theme(
    axis.text.x = element_text(face = "bold"),
    strip.background = element_blank(),
    strip.text.y = element_text(face = "bold")
    )
ggsave(paste0("figs/", "Weekday_mean_adj_dep_presentonly_insideman.png"), width=5, height=12, units="in", dpi=300)



# Inside small vessels
ggplot(data = wkly_means_ins_small,
       aes(x = Weekday, y = mean_wkdy_adj)) +
  geom_line(aes(group = Dep_ID),
            color = "gray70")+
  geom_pointrange(aes(ymin = sd_lower,
                      ymax = sd_upper,
                      group = npz_id,
                      color = ifelse(sd_lower <= 0, 'firebrick','blue')),
                  size = 0.25,
                  show.legend = FALSE) +
  geom_hline(yintercept = 0, 
             linetype = 3, # dotted line
             colour = "black") +
  scale_color_manual(values = c("red","black"))+
  labs(title = "Inside small vessels",
    y = "Mean-adjusted average N vessels / weekday",
       x = "Weekday")+
  facet_nested(rows = vars(network, npz_id, Dep_ID), 
               nest_line = element_line(),
               scales = "free_y") +
  theme(
    axis.text.x = element_text(face = "bold"),
    strip.background = element_blank(),
    strip.text.y = element_text(face = "bold")
    )
ggsave(paste0("figs/", "Weekday_mean_adj_dep_presentonly_inside_small.png"), width=5, height=12, units="in", dpi=300)



```



### Daily mean-adjusted diel presence

```{r mean adj diel}
# start by getting hourly counts
all_ves_diel <- all_sites_hp |>
  group_by(network, npz_id, Dep_ID, Dep, Begin_Date_loc, Begin_Hour_loc)|>
  # get daily vessel counts per weekday: total, inside, inside_small, maneuver, maneuver_small
  summarise(Total_ves_dep = sum(Total_Vessels, na.rm = TRUE),
            Total_inside = sum(total_inside, na.rm = TRUE),
            Total_inside_small = sum(total_inside_small, na.rm = TRUE),
            Inside_maneuver = sum(man_inside, na.rm = TRUE),
            Inside_maneuver_small = sum(man_inside_small, na.rm = TRUE)) |>
  mutate(Inside_transit = Total_inside - Inside_maneuver,
         Inside_transit_small = Total_inside_small - Inside_maneuver_small, 
         # Weekday = factor(Weekday, levels=c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday")),
         network = case_when(npz_id == "CGMP" | npz_id == "SIMP" ~ "Temperate East",
                             npz_id == "DMP" | npz_id == "NMP" ~ "Northwest", TRUE ~ "South-west"
         ),
         network = factor(network, levels = c("Temperate East","Northwest","South-west"))) 


# Calculate daily mean-adjusted counts per hour
daily_means_tot <- all_ves_diel |>
  # group to date level
  group_by(network, npz_id, Dep_ID, Dep, Begin_Date_loc) |>
  # get n_hours per day
  summarize(n_hrs = n(),
            # and total vessels per day
            day_sum = sum(Total_ves_dep, na.rm = TRUE)) |>
  # only keep full days n_hrs == 24
  # and days with vessels present
  filter(n_hrs == 24,
         day_sum >0) |>
  # join to get hourly counts back
  left_join(all_ves_diel) |>
  # re-group to daily level
  group_by(network, npz_id, Dep_ID, Dep, Begin_Date_loc) |>
  # calculate mean total presence per day with vessels
  mutate(mean_tot = mean(Total_ves_dep, na.rm = TRUE),
         # calculate mean-adjusted counts per hour
         tot_diel_adj = Total_ves_dep - mean_tot) |>
  # group to hour level per deployment
  group_by(network, npz_id, Dep_ID, Begin_Hour_loc) |>
  # get mean of mean-adjusted ounts per hour
  summarise(mean_diel_adj = mean(tot_diel_adj, na.rm =TRUE)) |>
  # get lower & upper bounds for SD around mean of adjusted counts
  mutate(sd_lower = mean_diel_adj - 0.5*(sd(mean_diel_adj)),
         sd_upper = mean_diel_adj + 0.5*(sd(mean_diel_adj)))
  


# Calculate daily mean-adjusted counts per hour
daily_means_ins <- all_ves_diel |>
  # group to date level
  group_by(network, npz_id, Dep_ID, Dep, Begin_Date_loc) |>
  # get n_hours per day
  summarize(n_hrs = n(),
            # and total vessels per day
            day_sum = sum(Total_inside, na.rm = TRUE)) |>
  # only keep full days n_hrs == 24
  # and days with vessels present
  filter(n_hrs == 24,
         day_sum > 0) |>
  # join to get hourly counts back
  left_join(all_ves_diel) |>
  # re-group to daily level
  group_by(network, npz_id, Dep_ID, Dep, Begin_Date_loc) |>
  # calculate mean total presence per day with vessels
  mutate(mean_ins = mean(Total_inside, na.rm = TRUE),
         # calculate mean-adjusted counts per hour
         ins_diel_adj = Total_inside - mean_ins) |>
  # group to hour level per deployment
  group_by(network, npz_id, Dep_ID, Begin_Hour_loc) |>
  # get mean of mean-adjusted ounts per hour
  summarise(mean_diel_adj = mean(ins_diel_adj, na.rm =TRUE)) |>
  # get lower & upper bounds for SD around mean of adjusted counts
  mutate(sd_lower = mean_diel_adj - 0.5*(sd(mean_diel_adj)),
         sd_upper = mean_diel_adj + 0.5*(sd(mean_diel_adj)))
  
            # mean_tot = mean(Total_ves_dep, na.rm = TRUE),
            # mean_ins = mean(Total_inside, na.rm = TRUE),
            # mean_ins_sm = mean(Total_inside_small, na.rm = TRUE),
            # mean_trans = mean(Inside_transit, na.rm = TRUE),
            # mean_trans_sm = mean(Inside_transit_small, na.rm = TRUE),
            # mean_man = mean(Inside_maneuver, na.rm = TRUE), 
            # mean_man_sm = mean(Inside_maneuver_small, na.rm = TRUE))
            
#   
# diel_mean_depid <- all_ves_diel |>
#   left_join(daily_means, by = c("Begin_Date_loc", "Dep"), keep = FALSE) |>
#   group_by(network.x, npz_id.x, Dep_ID.x, Begin_Hour_loc) |>
#   filter(n_hrs == 24, # only keep full days
#          # only keep days with presence for mean-adjusting
#          mean_tot > 0) |>
#   mutate(tot_diel_adj = Total_ves_dep - mean_tot,
#          ins_diel_adj = Total_inside - mean_ins,
#          ins_sm_adj = Total_inside_small - mean_ins_sm) |>
#   summarise(mean_diel = mean(tot_diel_adj, na.rm = TRUE),
#             mean_diel_ins = mean(ins_diel_adj, na.rm = TRUE),
#             mean_diel_ins_sm = mean(ins_sm_adj, na.rm = TRUE)) |>
#   mutate(sd_lower = mean_diel - 0.5*(sd(mean_diel)),
#          sd_upper = mean_diel + 0.5*(sd(mean_diel)),
#          sd_lower_ins = mean_diel_ins - 0.5*(sd(mean_diel_ins)),
#          sd_upper_ins = mean_diel_ins + 0.5*(sd(mean_diel_ins)),
#          sd_lower_ins_sm = mean_diel_ins - 0.5*(sd(mean_diel_ins)),
#          sd_upper_ins_sm = mean_diel_ins_sm + 0.5*(sd(mean_diel_ins_sm)))


# plot total vessels
ggplot(data = daily_means_tot,
       aes(x = Begin_Hour_loc, 
           y = mean_diel_adj)) +
  geom_line(aes(group = Dep_ID),
            color = "gray70")+
  geom_pointrange(aes(ymin = sd_lower,
                      ymax = sd_upper,
                      group = npz_id,
                      color = ifelse(sd_lower <= 0, 'firebrick','blue')),
                  size = 0.25,
                  show.legend = FALSE) +
  geom_hline(yintercept = 0, 
             linetype = 3, # dotted line
             colour = "black") +
  scale_color_manual(values = c("red","black"))+
  scale_x_continuous(breaks = seq(0,23,2))+
  labs(y = "Mean-adjusted average N vessels / hour",
       x = "Hour of Day (Local TZ)",
       title = "Total vessels")+
   facet_nested(rows = vars(network, npz_id, Dep_ID), 
               nest_line = element_line(),
               scales = "free_y") +
  theme(
    axis.text.x = element_text(face = "bold"),
    strip.background = element_blank(),
    strip.text.y = element_text(face = "bold")
    )
ggsave(paste0("figs/", "Diel_mean_adj_dep_presonly_withline.png"), width=8, height=12, units="in", dpi=300)



#### Combined total & inside to help visualize differences in patterns

ggplot(data = daily_means_tot,
       aes(x = Begin_Hour_loc, 
           y = mean_diel_adj)) +
  geom_hline(yintercept = 0, 
             linetype = 3, # dotted line
             colour = "black") +
  geom_line(aes(group = Dep_ID),
            color = "black")+
  geom_pointrange(aes(ymin = sd_lower,
                      ymax = sd_upper,
                      group = npz_id,
                      color = ifelse(sd_lower <= 0, 'firebrick','blue')),
                  size = 0.25,
                  show.legend = FALSE) +
  geom_line(data = daily_means_ins,
            aes(group = Dep_ID),
            color = "blue")+
  geom_pointrange(data = daily_means_ins,
                  aes(ymin = sd_lower,
                      ymax = sd_upper,
                      group = npz_id,
                      color = ifelse(sd_lower <= 0, 'salmon','chartreuse')),
                  size = 0.25,
                  show.legend = FALSE) +
  scale_color_viridis_d() + 
   scale_color_manual(values = c('black', # above 0, total
                                'blue', # above 0, inside
                                'black', # below 0, total
                                'blue' # below 0, inside
                                ))+
  scale_x_continuous(breaks = seq(0,23,2))+
  labs(y = "Mean-adjusted average N vessels / hour",
       x = "Hour of Day (Local TZ)",
       title = "Total vessels")+
   facet_nested(rows = vars(network, npz_id, Dep_ID), 
               nest_line = element_line(),
               scales = "free_y") +
  theme(
    axis.text.x = element_text(face = "bold"),
    strip.background = element_blank(),
    strip.text.y = element_text(face = "bold")
    )
ggsave(paste0("figs/", "Diel_mean_adj_dep_combined_1color.png"), width=8, height=12, units="in", dpi=300)






# Plot inside vessels
ggplot(data = daily_means_ins,
       aes(x = Begin_Hour_loc, 
           y = mean_diel_adj)) +
  geom_line(aes(group = Dep_ID),
            color = "gray70")+
  geom_pointrange(aes(ymin = sd_lower,
                      ymax = sd_upper,
                      group = npz_id,
                      color = ifelse(sd_lower <= 0, 'firebrick','blue')),
                  size = 0.25,
                  show.legend = FALSE) +
  geom_hline(yintercept = 0, 
             linetype = 3, # dotted line
             colour = "black") +
  scale_color_manual(values = c("red","black"))+
  scale_x_continuous(breaks = seq(0,23,2))+
  labs(y = "Mean-adjusted average N vessels / hour",
       x = "Hour of Day (Local TZ)",
       title = "Inside vessels")+
   facet_nested(rows = vars(network, npz_id, Dep_ID), 
               nest_line = element_line(),
               scales = "free_y") +
  theme(
    axis.text.x = element_text(face = "bold"),
    strip.background = element_blank(),
    strip.text.y = element_text(face = "bold")
    )
ggsave(paste0("figs/", "Diel_mean_adj_dep_presonly_ins.png"), width=8, height=12, units="in", dpi=300)



# # summarize to npz level
# diel_mean_adj_npz <- all_ves_diel |>
#   left_join(daily_means, by = c("Begin_Date_loc", "Dep"), keep = FALSE) |>
#   group_by(network.x, npz_id.x, Begin_Hour_loc) |>
#   mutate(tot_diel_adj = Total_ves_dep - mean_tot,
#          ins_diel_adj = Total_inside - mean_ins,
#          ins_sm_adj = Total_inside_small - mean_ins_sm) |>
#   summarise(mean_diel = mean(tot_diel_adj, na.rm = TRUE),
#             mean_diel_ins = mean(ins_diel_adj, na.rm = TRUE),
#             mean_diel_ins_sm = mean(ins_sm_adj, na.rm = TRUE)) |>
#   mutate(sd_lower = mean_diel - 0.5*(sd(mean_diel)),
#          sd_upper = mean_diel + 0.5*(sd(mean_diel)),
#          sd_lower_ins = mean_diel_ins - 0.5*(sd(mean_diel_ins)),
#          sd_upper_ins = mean_diel_ins + 0.5*(sd(mean_diel_ins)),
#          sd_lower_ins_sm = mean_diel_ins - 0.5*(sd(mean_diel_ins)),
#          sd_upper_ins_sm = mean_diel_ins_sm + 0.5*(sd(mean_diel_ins_sm)))
#    
#    
# ggplot(data = diel_mean_npz,
#        aes(x = factor(Begin_Hour_loc), 
#            y = mean_diel)) +
#   geom_hline(yintercept = 0, colour = "grey50") +
#   geom_pointrange(aes(ymin = sd_lower,
#                       ymax = sd_upper,
#                       group = npz_id.x,
#                       color = ifelse(sd_lower <= 0, 'firebrick','blue')),
#                   size = 0.25,
#                   show.legend = FALSE) +
#   geom_hline(yintercept = 0, 
#              linetype = 3, # dotted line
#              colour = "black") +
#   scale_color_manual(values = c("red","black"))+
#   labs(y = "Mean-adjusted average N vessels / hour",
#        x = "Hour of Day (Local TZ)")+
#    facet_nested(rows = vars(network.x, npz_id.x), 
#                nest_line = element_line(),
#                scales = "free_y") +
#   theme(
#     axis.text.x = element_text(face = "bold"),
#     strip.background = element_blank(),
#     strip.text.y = element_text(face = "bold")
#     )
# 
# ggsave(paste0("figs/", "Diel_mean_adj_npz_presonly.png"), width=8, height=10, units="in", dpi=300)


# Inside small vessels
# 
# ggplot(data = diel_mean_npz |> filter(npz_id.x != "SWCMP"),
#        aes(x = factor(Begin_Hour_loc), 
#            y = mean_diel_ins_sm)) +
#   geom_hline(yintercept = 0, colour = "grey50") +
#   geom_pointrange(aes(ymin = sd_lower_ins_sm,
#                       ymax = sd_upper_ins_sm,
#                       group = npz_id.x,
#                       color = ifelse(mean_diel_ins_sm < 0, 'firebrick','blue')),
#                   size = 0.25,
#                   show.legend = FALSE) +
#   scale_color_viridis_d(begin = 0.3, end = 0.7, 
#                         option = "H",
#                         direction = -1)+
#   labs(y = "Mean-adjusted average N vessels / hour",
#        x = "Hour of Day (Local TZ)")+
#   facet_grid(rows = vars(npz_id.x)) +
#   theme(
#     axis.text.x = element_text(face = "bold"),
#     strip.text.y = element_text(face = "bold")
#     )
# 
# ggsave(paste0("figs/", "Diel_mean_adj_npz_ins_sm.png"), width=8, height=10, units="in", dpi=300)








```



```{r pooling SIMP sites}
### Compare CGMP deployments
cgmp_wkdy <- all_ves_wkdy |>
  filter(Site_ID == "CG") |>
  group_by(Dep, Weekday) |>
  summarise(Total_ves_dep = sum(Total_ves_dep)) |>
  pivot_wider(names_from = Dep, values_from = Total_ves_dep)

### Compare SIMP deployments
simp_wkdy <- all_ves_wkdy |>
  filter(npz_id == "SIMP")|>
  group_by(Dep, Weekday) |>
  summarise(Total_ves_dep = sum(Total_ves_dep)) |>
  pivot_wider(names_from = Dep, values_from = Total_ves_dep)

# Chi squared tests
# EP vs. WP in same deployment --> can we pool them?

chisq.test(as.data.frame(simp_wkdy[,2:3], row.names = FALSE)) # NS
chisq.test(as.data.frame(simp_wkdy[,4:5], row.names = FALSE)) # NS
chisq.test(as.data.frame(simp_wkdy[,6:7], row.names = FALSE)) # sig (but NS for same dates)

# test if WP subset to same dates is different
simp_epwp_dates <- all_ves_wkdy |>
  filter(npz_id == "SIMP", 
         Begin_Date_loc > "2019-06-10" & Begin_Date_loc < "2019-06-30") |>
  group_by(Dep, Weekday) |>
  summarise(Total_ves_dep = sum(Total_ves_dep)) |>
  pivot_wider(names_from = Dep, values_from = Total_ves_dep)


```


## Daily proportional hours present

Summarize by weekday

```{r prop across weekdays}

# from full hp dataset
# group down to date level to get daily vessel counts for each category (total, inside, inside_small, transit, maneuver)
# remove any days without vessels present
# group by deployment and weekday
# get proportion of vessel counts for each category occurring on each weekday
# 

all_ves_wkdy <- all_sites_hp |>
  group_by(network, npz_id, Site_ID, Dep, Weekday)|>
  # get daily vessel countsper weekday: total, inside, inside_small, maneuver, maneuver_small
  summarise(Total_ves_dep = sum(Total_Vessels, na.rm = TRUE),
            Total_inside = sum(total_inside, na.rm = TRUE),
            Total_inside_small = sum(total_inside_small, na.rm = TRUE),
            Inside_maneuver = sum(man_inside, na.rm = TRUE),
            Inside_maneuver_small = sum(man_inside_small, na.rm = TRUE)) |>
  mutate(Inside_transit = Total_inside - Inside_maneuver,
         Inside_transit_small = Total_inside_small - Inside_maneuver_small, 
         Weekday = factor(Weekday, levels=c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday")),
         network = case_when(Site_ID == "CG" | Site_ID == "EP" | Site_ID == "WP" ~ "Temperate East",
                             Site_ID == "DNE" | Site_ID == "DSW" | Site_ID == "NGN"| Site_ID == "NGS" ~ "Northwest", TRUE ~ "South-west"
         ),
         network = factor(network, levels = c("Temperate East","Northwest","South-west"))) |>
  filter(Total_ves_dep > 0) 

levels(all_ves_wkdy$Weekday) <- c("M","T","W","R","F","Sa","Su")


# function to calculate Agresti-Coull interval
# from https://towardsdatascience.com/five-confidence-intervals-for-proportions-that-you-should-know-about-7ff5484c024f
AgrestiInterval <- function(x, n, conf.level = 0.95){
 p <- (x+2)/(n+4)
 sd <- sqrt(p*((1-p)/(n+4)))
 z <- qnorm(c( (1 - conf.level)/2, 1- (1-conf.level)/2)) #returns the value of thresholds at which conf.level has to be cut at. for 95% CI, this is -1.96 and +1.96
 ci <- p + z*sd
 return(ci)
 }

wkdy_prop_table <- all_ves_wkdy |>
  group_by(network, npz_id, Site_ID, Dep) |>
  mutate(Inside_transit = Total_inside - Inside_maneuver,
         Inside_transit_small = Total_inside_small - Inside_maneuver_small,
         tot_prop = Total_ves_dep/sum(Total_ves_dep),
         ins_prop = Total_inside/sum(Total_inside),
         small_prop = Total_inside_small/sum(Total_inside_small),
         ins_trans_prop = Inside_transit/sum(Inside_transit),
         small_trans_prop = Inside_transit_small/sum(Inside_transit_small),
         ins_man_prop = Inside_maneuver/sum(Inside_maneuver),
         small_man_prop = Inside_maneuver_small/sum(Inside_maneuver_small))

wkdy_prop_long <- wkdy_prop_table |>
  pivot_longer(cols = -c(network, npz_id, Site_ID, Dep, Weekday), 
               values_to = "Weekday_prop", names_to = "Ves_behav") |>
  filter(grepl("prop", Ves_behav)) |>
  # relevel categorical values for plotting
  mutate(Ves_behav = fct(Ves_behav),
         Ves_behav = fct_relevel(Ves_behav, c("tot_prop","small_prop", "ins_prop", "small_trans_prop","small_man_prop", "ins_trans_prop","ins_man_prop")),
         # network = fct(network),
         network = fct_relevel(network, c("Temperate East","Northwest")), 
         npz_id = fct(npz_id),
         npz_id = fct_relevel(npz_id, c("CGMP","SIMP","DMP","NMP")))



# TO DO:
# - calculate A-C confidence intervals for each proportion
# - add CI's to plot
```




## Proportional hours present

For days with presence AND 24 total hours (i.e., full days only)
- Calculate proportion of hours present (n/24)
- Summarize by weekday (median, 25th - 75th percentiles)


```{r weekday prophrs}
#### Proportional by hours present rather than counts ####

# start with full hp dataset
all_perc_weekday <- all_sites_hp |>
  # group by date to get % presence per day 
  group_by(network, npz_id, Site_ID, Dep, Date_noyr, Begin_Date_loc, Weekday, ves_yn) |>
  # get total number of hours for Y and N groups per each hour per deployment
  summarize(n_hours = n())|>
  # only keep full days
  filter(sum(n_hours) == 24) |>
  # calculate proportion hours present per day
  mutate(freq = n_hours/sum(n_hours)) |>
  # ungroup to fill in any missing values for columns
  ungroup() |>
  complete(Dep, Weekday, ves_yn,
           fill = list(ves_yn = "Y", freq = 0)) |>
  # regroup
  group_by(network, npz_id, Site_ID, Dep, Begin_Date_loc, Weekday, ves_yn) |>
  # reshape for easier plotting
  pivot_longer(cols = c("freq"),
               values_to = "Perc_per_hour") |>
  mutate(Weekday = factor(Weekday, levels=c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday")),
          network = case_when(Site_ID == "CG" | Site_ID == "EP" | Site_ID == "WP" ~ "Temperate East",
                              Site_ID == "DNE" | Site_ID == "DSW" | Site_ID == "NGN"| Site_ID == "NGS" ~ "Northwest",
                              TRUE ~ "South-west"
          ),
          network = factor(network, levels = c("Temperate East","Northwest","South-west")))
levels(all_perc_weekday$Weekday) <- c("M","T","W","R","F","Sa","Su")


# summarize proportion hours present

# by deployment
perc_wkdy_summ_dep <- all_perc_weekday |>
  # only interested in proportion present
  filter(ves_yn == "Y") |>
  ungroup()|>
  # group down to weekday level to summarize
  group_by(network, npz_id, Site_ID, Dep, Weekday)|>
  summarise(median_prop = median(Perc_per_hour),
            qt25 = quantile(Perc_per_hour, probs = 0.25),
            qt75 = quantile(Perc_per_hour, probs = 0.75),
            min_perc = min(Perc_per_hour, na.rm = TRUE),
            max_perc = max(Perc_per_hour, na.rm = TRUE))

# by npz
perc_wkdy_summ_npz <- all_perc_weekday |>
  filter(ves_yn == "Y") |>
  ungroup()|>
  group_by(network, npz_id, Weekday)|>
  summarise(median_prop = median(Perc_per_hour),
            qt25 = quantile(Perc_per_hour, probs = 0.25),
            qt75 = quantile(Perc_per_hour, probs = 0.75),
            min_perc = min(Perc_per_hour, na.rm = TRUE),
            max_perc = max(Perc_per_hour, na.rm = TRUE))

```


Daily proportion of hours present, summarized by weekday
"For days with vessels present, what proportion of the day had vessels?
When those are summarized by Weekday, what's the median value for each weekday at each NPZ?"

```{r weekday proportion hours}

## Try individual npz's for a single network

## Temperate East
ggplot(data = all_perc_weekday |>
         filter(ves_yn == "Y",
                !is.na(npz_id),
                network == "Temperate East"),
       aes(x = Weekday,
           y = Perc_per_hour,
           fill = npz_id,
           color = npz_id))+
  geom_boxplot(
    position = position_dodge2(preserve = "single"),
    alpha = 0.75,
    width = 0.5)+
  scale_fill_viridis_d(end = 0.75,
                       direction = -1
  )+
  scale_color_viridis_d(end = 0.75,
                        direction = -1,
                        guide = "none")+
  facet_grid(rows = vars(Dep),
             drop = TRUE) +
  labs(x = "Weekday",
       y = "Proportion hours per weekday",
       fill = "Deployment") +
  theme(
    strip.text.y = element_text(face = "bold"),
    axis.text.y = element_text(face = "bold"),
    axis.text.x = element_text(face = "bold")
  )
# 
ggsave(paste0("figs/", "Weekday_prop_hrs_npz_tempeast_dep.png"), width=6, height=10,
       units="in", dpi=300)




## Northwest
ggplot(data = all_perc_weekday |>
         filter(ves_yn == "Y",
                !is.na(npz_id),
                network == "Northwest"),
       aes(x = Weekday,
           y = Perc_per_hour,
           fill = npz_id,
           color = npz_id))+
  geom_boxplot(
    position = position_dodge2(preserve = "single"),
    alpha = 0.75,
    width = 0.5)+
  scale_fill_viridis_d(end = 0.75,
                       direction = -1
  )+
  scale_color_viridis_d(end = 0.75,
                        direction = -1,
                        guide = "none")+
  facet_grid(rows = vars(npz_id),
             drop = TRUE) +
  labs(x = "Weekday",
       y = "Proportion hours per weekday",
       fill = "NPZ") +
  theme(
    strip.text.y = element_text(face = "bold"),
    axis.text.y = element_text(face = "bold"),
    axis.text.x = element_text(face = "bold")
  )
# 
ggsave(paste0("figs/", "Weekday_prop_hrs_npz_northwest.png"), width=6, height=10,
       units="in", dpi=300)

## South-west
ggplot(data = all_perc_weekday |>
         filter(ves_yn == "Y",
                !is.na(npz_id),
                network == "South-west"),
       aes(x = Weekday,
           y = Perc_per_hour,
           fill = npz_id,
           color = npz_id))+
  geom_boxplot(
    position = position_dodge2(preserve = "single"),
    alpha = 0.75,
    width = 0.5)+
  scale_fill_viridis_d(end = 0.75,
                       direction = -1
  )+
  scale_color_viridis_d(end = 0.75,
                        direction = -1,
                        guide = "none")+
  facet_grid(rows = vars(npz_id),
             drop = TRUE) +
  labs(x = "Weekday",
       y = "Proportion hours per weekday",
       fill = "NPZ") +
  theme(
    strip.text.y = element_text(face = "bold"),
    axis.text.y = element_text(face = "bold"),
    axis.text.x = element_text(face = "bold")
  )
# 
ggsave(paste0("figs/", "Weekday_prop_hrs_npz_southwest.png"), width=6, height=10,
       units="in", dpi=300)


# no facets, plotting at network level

ggplot(data = all_perc_weekday |>
         filter(ves_yn == "Y",
                Perc_per_hour > 0, 
                !is.na(npz_id)),
       aes(x = Weekday,
           y = Perc_per_hour,
           fill = network,
           color = network))+
  geom_boxplot(
    position = position_dodge2(preserve = "single"),
    alpha = 0.75,
    width = 0.5)+
  scale_fill_viridis_d(end = 0.75,
                       direction = -1
  )+
  scale_color_viridis_d(end = 0.75,
                        direction = -1,
                        guide = "none")+
  # facet_grid(rows = vars(npz_id),
  #            drop = TRUE) +
  labs(x = "Weekday",
       y = "Hours present / hours per day",
       fill = "Network") +
  theme(
    strip.text.y = element_text(face = "bold"),
    axis.text.y = element_text(face = "bold"),
    axis.text.x = element_text(face = "bold")
  )
ggsave(paste0("figs/", "Weekday_prop_hrs_nofacets.png"), width=6, height=10,
       units="in", dpi=300)





```



## Plotting weekday proportions

```{r weekday plots}


wkday_prop_total <- wkdy_prop_long |>
  filter(Ves_behav == "tot_prop")

ggplot(data = wkday_prop_total,
       aes(x = Weekday, 
           y = Weekday_prop)) +
  geom_col(aes(fill = network),
           position = position_dodge()) +
  scale_fill_viridis_d(end = 0.75, direction = -1)+
  labs(y = "Proportion", 
       x = "Weekday", 
       fill = "Network")+  
  facet_grid(rows = vars(npz_id)) +
   theme(
    axis.text.x = element_text(face = "bold"),
    strip.text.y = element_text(face = "bold")
    )



ggsave(paste0("figs/", "Weekday_props_npz.png"), width=8, height=10,
       units="in", dpi=300)




wkday_prop_ins <- wkdy_prop_long |>
  filter(Ves_behav == "ins_prop" | Ves_behav == "ins_trans_prop" | Ves_behav == "ins_man_prop")

ggplot(data = wkday_prop_ins |> filter (npz_id != "MMP" & npz_id != "SWCMP"),
       aes(x = Weekday, 
           y = Weekday_prop)) +
  geom_col(aes(fill = Ves_behav),
           position = position_dodge()) +
  scale_fill_viridis_d(end = 0.75, direction = -1)+
  labs(y = "Proportion", 
       x = "Weekday", 
       fill = "Network")+  
  facet_grid(rows = vars(npz_id),
             cols = vars(Ves_behav)) +
   theme(
    axis.text.x = element_text(face = "bold"),
    strip.text.y = element_text(face = "bold")
    )



ggsave(paste0("figs/", "Weekday_props_npz_inside.png"), width=8, height=10, units="in", dpi=300)






```



### Export CSV summary tables

```{r export tables}

# Total vessel presence
write_csv(all_ves_by_site, 
          file = "output/All_ves_by_dep.csv")
write_csv(all_ves_by_npz, 
          file = "output/All_ves_by_npz.csv")
write_csv(all_ves_wkdy, 
          file = "output/All_ves_by_weekday.csv")
write_csv(cgmp_wkdy,
          file = "output/CGMP_by_weekday.csv")

write_csv(all_perc_weekday, 
          file = "output/All_ves_prophr_weekday.csv")
write_csv(perc_wkdy_summ_npz, 
          file = "output/All_ves_prophr_weekday_npz.csv")
write_csv(perc_wkdy_summ_dep, 
          file = "output/All_ves_prophr_weekday_dep.csv")
write_csv(all_ves_diel, 
          file = "output/All_ves_diel_counts.csv")
write_csv(diel_mean_adj, 
          file = "output/All_ves_diel_mean_adj.csv")

```
